---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# epirice

A fork of the R [cropsim package](https://r-forge.r-project.org/R/?group_id=335) designed to make using the EPIRICE model for rice diseases easier

The original manuscript, Savary et al. (2012), which details the model and results of its use to model global epidemics of rice diseases was published in Crop Protection,

> Serge Savary, Andrew Nelson, Laetitia Willocquet, Ireneo Pangga and Jorrel Aunario. Modeling and mapping potential epidemics of rice diseases globally. Crop Protection, Volume 34, 2012, Pages 6-17, ISSN 0261-2194 DOI: [http://dx.doi.org/10.1016/j.cropro.2011.11.009](http://dx.doi.org/10.1016/j.cropro.2011.11.009).
  
# Installation

`epirice` is not yet on CRAN.
You can install it this way.

```{r install, eval=FALSE}
if (!require("remotes"))
  install.packages("remotes")
remotes::install_github("adamshsparks/epirice"
)
```

# Meta

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
