---
title: "epirice"
author: "Adam H Sparks"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{epirice}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{ggplot2}
  %\VignetteDepends{data.table}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 7, fig.height = 7)
```

## Introduction to epirice

`epirice` provides an R package of the EPIRICE model as described in Savary *et al.* 2012. Default values derived from the literature suitable for modelling unmanaged disease severity of five rice diseases,

* bacterial blight,

* brown spot,

* leaf blast,

* sheath blight and

* tungro
are provided.
The model uses daily weather data to estimate disease severity.
A function, `get_wth()` is provided to simplify downloading weather data from the [https://power.larc.nasa.gov](NASA POWER web API) that are suitable for use in `epirice`.

Using the package functions is designed to be straightforward for modelling rice disease risks, but flexible enough to accommodate other pathosystems using the `SEIR()` function.
If you are interested in modelling other pathosystems, please refer to Savary *et al.* 2012 for the development of the parameters that were used for the rice diseases as derived from the existing literature and are implemented in the individual disease model functions.

### Getting started

Load the library.

```{r epirice}
library(epirice)
```

### Get weather data

The most simple way to use the model is to download weather data from NASA POWER using `get_wth()`, which provides the data in a format suitable for use in the model and is freely available.

```{r power}
# Fetch weather for year 2000 wet season at the IRRI Zeigler Experiment Station
 wth <- get_wth(
   lonlat = c(121.25562, 14.6774),
   dates = c("2000-06-30", "2000-12-31")
 )

wth
```

### Predict bacterial blight

All of the `predict_()` family of functions work in exactly the same manner.
You provide them with weather data and an emergence date, that falls within the weather data provided, and they will return a data frame of disease severity over the season and other values associated with the model.
See `?SEIR()` for more on the values returned.

```{r predict_bb}
 bb <- predict_bacterial_blight(wth, emergence = "2000-07-01")
```

```{r bb}
summary(bb)
```

## Plotting using ggplot2

The data are in a wide format by default and need to be converted to long format for use in `ggplot2` if you wish to plot more than one variable at a time.

```{r ggplot2}
library(ggplot2)
library(tidyr)
```

### Sites

The model records the number of sites for each bin daily, this can be graphed as follows.

```{r plot_sites}
dat <- pivot_longer(
  bb,
  cols = c("diseased", "removed", "latent", "infectious"),
  names_to = "site",
  values_to = "value"
)

ggplot(data = dat,
       aes(
         x = dates,
         y = value,
         shape = site,
         linetype = site
       )) +
  labs(y = "Sites",
       x = "Date") +
  geom_line(aes(group = site, colour = site)) +
  geom_point(aes(colour = site)) +
  labs(title = "Bacterial blight site states over time",
       caption = "Data source: NASA POWER API")
```

### Severity

Plotting severity over time does not require any data manipulation.

```{r plot_severity}
ggplot(data = dat,
       aes(x = dates,
           y = severity)) +
  labs(y = "Severity",
       x = "Date") +
  geom_line() +
  geom_point() +
  labs(title = "Bacterial blight disease progress over time",
       caption = "Data source: NASA POWER API")
```

# References
Serge Savary, Andrew Nelson, Laetitia Willocquet, Ireneo Pangga and Jorrel Aunario. Modeling and mapping potential epidemics of rice diseases globally. *Crop Protection*, Volume 34, 2012, Pages 6-17, ISSN 0261-2194 DOI: [10.1016/j.cropro.2011.11.009](https://doi.org/10.1016/j.cropro.2011.11.009).
